\thesistitle

### 2019-08-22 sections/background.tex
Background Ana

\chapter{Background}
\label{ch:background}
This chapter will present the necessary background information for this thesis. Here, we define some basic terminology that will be used throughout this thesis.

\section{Code clone terminology}\label{sec:terminology}
Many studies present different definitions for code clone concepts. For this study, we mainly use the definitions from Bruntink et al. \cite{bruntink2005use}, Roy et al. \cite{roy2007survey} and Jiang et al. \cite{jiang2007deckard}. We created a summary of these concepts, which can be found in Table~\ref{tab:clone-terminology}. The terms displayed in this table show how tokens in code map to clones.

@@

This example can be represented as a tree structure using the concepts from Table~\ref{tab:clone-terminology} in Figure~\ref{fig:terminologyexample}.

@@

\section{Refactoring methods}
In Martin Fowler's popular ``Refactoring'' book \cite{fowler2018refactoring}, he exclaims that \textit{``if you see the same code structure in more than one place, you can be sure that your program will be better if you find a way to unify them''}. He describes several methods to deal with duplication, dependent on the context of the clone. In this section we describe the methods which we use in this study.

\subsection{Extract Method}
Current clone refactoring research shows that the ``Extract Method'' refactoring technique can refactor most duplication problems \cite{fontana2015duplicated, tsantalis2015assessing, white2016deep}. Method extraction moves matching functionality in method bodies to a common place, namely a new method \cite{fowler2018refactoring}. An example of this procedure is displayed in figure \cite{fig:extractmethod}.

@@

\subsection{Move method}
In the example of the previous section, both duplicated parts are in the same method. However, a study by Fontana et al. \cite{fontana2015duplicated} shows that this is most often not the case. Based on the relation between clone instances, the extracted method must be moved to be accessible by all locations of the clone instances. This might require different techniques.

\subsubsection{Pull up method}
A refactoring technique to move a method up in its inheritance structure is called ``Pull up method'' \cite{fowler2018refactoring}. This way, if cloned methods are related in any way through inheritance, they can be called by both classes by placing the method in a class they both have in common. This way its possible to refactor both fully cloned methods (by just pulling up the method) and partially cloned methods (by first performing method extraction and then pulling up the refactored method).

\subsubsection{Create class abstraction based on implicit relations}
Duplication in source code is an implicit relation between fragments of source code. If two classes have many of these implicit relations, then the implementation should be refactored to make this relation explicit. If the classes do not yet have a parent/super class, a parent class can be created and the common functionality can be placed in this newly created class. This makes the relation between these classes explicit and reduces duplication.

\subsubsection{Providing default implementations for common functionality}
Looking specifically at the Java programming language, recent versions of Java introduce a new method of reducing duplication through refactoring. Java has the concept of \textit{interfaces} as an abstract type that is used to specify a behavior that classes must implement. Since Java version 8, Java also supports default implementations to be provided in interfaces \cite{mohnen2002interfaces}. This gives an opportunity to make the relation between classes explicit and reduce duplication. It can be used in instances where creating a new parent class for duplicated classes is undesirable.

@@

\textit{External classes} are classes that a software system uses, but do not belong to this specific software system. Typically, the source code of external classes are not included in the source code of the project. Most often these external dependencies are referenced in a file that its build automation system uses to fetch a projects' dependencies.

Regarding refactoring, a software system can often not change the source code of its dependencies. This can sometimes be a burden, when an external dependency does not use proper abstractions that are required by a software system. Because of that, in this study, we differentiate between internal and external classes to see what impact they have on the refactoring process.\todo{you don't consider external right?}

\subsection{Maven}
For this study we perform a shallow analysis on external dependencies, in order to derive more context for internally used concepts. As these external dependencies are most often not included in a software systems' source code, we must first use the projects' build automation system to gather the dependencies. To limit the scope of this process, we decided to focus on only the \textit{Maven} build automation system.

>>>>

\chapter{Background}
\label{ch:background}
In this chapter we define some basic terminology that is used throughout this thesis.

\section{Code clone terminology}\label{sec:terminology}
Many studies present different definitions for code clone concepts. For this study, we mainly use the definitions from Bruntink et al. \cite{bruntink2005use}, Roy et al. \cite{roy2007survey} and Jiang et al. \cite{jiang2007deckard}. A summary of these concepts can be found in Table~\ref{tab:clone-terminology}. The terms displayed in this table show how tokens in code map to clones.

@@

This example can be represented as a tree structure, as shown in Figure~\ref{fig:terminologyexample}.

@@

\section{Refactoring techniques}
In Martin Fowler's popular ``Refactoring'' book \cite{fowler2018refactoring}, he exclaims that \textit{``if you see the same code structure in more than one place, you can be sure that your program will be better if you find a way to unify them''}. He describes several techniques to deal with duplication, dependent on the context of the clone. In this section we describe the methods that we use in this study.

\subsection{Extract Method}
Current clone refactoring research shows that the ``Extract Method'' refactoring technique can refactor most duplication problems \cite{fontana2015duplicated, tsantalis2015assessing, white2016deep}. ``Extract Method'' moves matching functionality in method bodies to a common place, namely a new method \cite{fowler2018refactoring}. An example of this procedure is displayed in figure \ref{fig:extractmethod}.

@@

\subsection{Move method}
In the example of the previous section, both duplicated parts are in the same method. However, a study by Fontana et al. \cite{fontana2015duplicated} shows that this is most often not the case. Based on the relation between clone instances, the extracted method must be moved to be accessible by all locations of the clone instances. This can require different techniques.

\subsubsection{Pull up method}
A refactoring technique to move a method up in its inheritance structure is called ``Pull up method'' \cite{fowler2018refactoring}. This way, if cloned methods are related in any way through inheritance, they can be called by both classes by placing the method in a class they both have in common. This way its possible to refactor both fully cloned methods (by just pulling up the method) and partially cloned methods (by first performing method extraction and then pulling up the refactored method). Anexample of this refactoring method is shown in Figure~\ref{fig:pullupmethod}.

\begin{figure}[H]
\begin{parcolumns}{2}
\colchunk[1]{
\begin{javacode}
// Original
class Vehicle{
  void start(){
|\highlightYellow|    putKeyIn();
|\highlightYellow|    turnKey();
    drive();
  }
}

class Car extends Vehicle{
  void startEngine(){
|\highlightYellow|    putKeyIn();
|\highlightYellow|    turnKey();
  }
}
\end{javacode}}
\colchunk[2]{
\begin{javacode}
// Refactored
class Vehicle{
  void start(){
|\highlightYellow|    startEngine();
    drive();
  }

  void startEngine(){
|\highlightYellow|    putKeyIn();
|\highlightYellow|    turnKey();
  }
}

class Car extends Vehicle{
}
\end{javacode}}
\end{parcolumns}
\caption{Refactoring a clone class using ``Pull Up Method''.}
\label{fig:pullupmethod}
\end{figure}

\subsubsection{Create class abstraction based on implicit relations}
Duplication in source code is an implicit relation between fragments of source code. If two classes have many of these implicit relations, then the implementation should be refactored to make this relation explicit \cite{fowler2018refactoring}. If the classes do not yet have a parent/super class, a parent class can be created and the common functionality can be placed in this newly created class. This makes the relation between these classes explicit and reduces duplication.

\begin{figure}[H]
\begin{parcolumns}{2}
\colchunk[1]{
\begin{javacode}
// Original
class Truck{
  void startEngine(){
|\highlightYellow|    putKeyIn();
|\highlightYellow|    turnKey();
  }
}

class Car{
  void startEngine(){
|\highlightYellow|    putKeyIn();
|\highlightYellow|    turnKey();
  }
}
\end{javacode}}
\colchunk[2]{
\begin{javacode}
// Refactored
class Vehicle{
  void startEngine(){
|\highlightYellow|    putKeyIn();
|\highlightYellow|    turnKey();
  }
}

class Truck extends Vehicle{
}

class Car extends Vehicle{
}
\end{javacode}}
\end{parcolumns}
\caption{Refactoring a clone class using ``Create class abstraction''.}
\label{fig:createclassabstraction}
\end{figure}

\subsubsection{Providing default implementations for common functionality}
In Java, C\#, Python and possibly other object-oriented languages, the programmer can create interfaces of which a class can implement any number. Such interfaces can (in Java, C\# and Python) provide default implementations for common functionality \cite{mohnen2002interfaces}. This gives an opportunity to make the relation between classes explicit and reduce duplication. It can be used in instances where creating a new parent class for duplicated classes is undesirable. An example of this is shown in Figure~\ref{fig:createinterfaceabstraction}, in which two unrelated types (Human and Radio) share a common property (making sound).

\begin{figure}[H]
\begin{parcolumns}{2}
\colchunk[1]{
\begin{javacode}
// Original
class Human implements MakesSound{
  void broadcastNews(){
|\highlightYellow|    sayNews();
|\highlightYellow|    sayWeather();
  }
}

class Radio implements MakesSound{
  void broadcastNews(){
|\highlightYellow|    sayNews();
|\highlightYellow|    sayWeather();
  }
}

interface MakesSound{
  void sayNews();
  void sayWeather();
}
\end{javacode}}
\colchunk[2]{
\begin{javacode}
// Refactored
class Human implements MakesSound{
}

class Radio implements MakesSound{
}

interface MakesSound{
  default void broadcastNews(){
|\highlightYellow|    sayNews();
|\highlightYellow|    sayWeather();
  }

  void sayNews();
  void sayWeather();
}
\end{javacode}}
\end{parcolumns}
\caption{Refactoring a clone class creating a default implementation for common functionality.}
\label{fig:createinterfaceabstraction}
\end{figure}

@@

\textit{External classes} are classes that a software system uses but do not belong to this specific software system (but to an external dependency). Most often these external dependencies are referenced in a file that its build automation system uses to fetch a projects' dependencies.

Regarding refactoring, a software system can often not change the source code of its dependencies. This can sometimes be a burden, when an external dependency does not use proper abstractions that are required by a software system. In this study we take external classes into consideration when choosing a refactoring technique, to be sure not to modify them.

\subsection{Maven}
For this study we perform a shallow analysis on external dependencies, in order to derive more context for internally used concepts. As these external dependencies are most often not included in a software systems' source code, we must first use the projects' build automation system to gather the dependencies. To limit the scope of this process, we decided to focus on only the \textit{Maven}\footnote{Apache Maven is a build automation system mainly used for Java: \url{https://maven.apache.org/}} build automation system.
